Imputation of Incomplete Multilevel Data with \proglang{R}

### 2023-11-24 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
Update imputation workflow

### Main analysis
When imputing clustered data, our first consideration should be the research question and the type of analysis that researchers intend to perform, assuming that non-incomplete values were present. This initial assessment provides valuable insights into the data's structure (see the level table), relationships among variables, and the causes of missingness.

When multiple imputation is used to deal with missing data, as the imputation and analysis process is performed separately, it is necessary that imputation model being congenial with the main analysis model (Meng, 1994), e.g. if the main model accounts for the hierarchical structure also imputation model should do it (Audigier, 2021).

### Exploration of available data
Then it is required to explore available information across clusters. This can involve evaluating the Intra-class Correlation Coefficient (ICC) to determine if a one-level imputation method is sufficient, particularly if there are no substantial differences between clusters. Additionally, it can be done by visualizing the data using available package as **ggmice**, for instance examining density plots can help us understand the variable type (normal, Poisson, categorical) and its range of plausible values. Exploring relationships between covariates through bivariate plots can be valuable in identifying potential interactions or non-linearities.

In the context of missingness, we should assess patterns of systematic and sporadic missing values and explore the relationships between the missingness of an incomplete variable and observable data. This evaluation allows us to consider the plausibility of Missing at Random (MAR) mechanisms over Missing Completely at Random (MCAR) methods and helps in selecting prediction variables to be included in the imputation model (considering overflux).

### Selecting plausible imputation methods
Before to implement multiple imputation it is requiered to evaluate if simpler imputation methods are suffice: 
- For example, certain incomplete variables may not necessitate stochastic imputation methods like MICE. Instead, they can be effectively addressed through deductive imputation, where incomplete values can be inferred from logic and deterministic relationships between variables. This approach is particularly useful for variables that are functions of others, e.g. a person's BMI from their weight and height, determining values for one-level variables from two-level ones, e.g. in the context of Individual Participant Data (IPD), incomplete information can be extracted from metadata, such as deducing incomplete data about abortion in a country where abortion is illegal, or engaging in cross-temporal or protocol's deduction, like imputing missing test values for deceased patients.

- Under monotonic pattern, the use of regression models
- CC when the proportion of missing is minimal <5%.
- Analysis that are robust to missing data, for instance FIML models such as mixed model do not require imputation models when only the outcome variable is missing.An alternative approach to missing data is to use Full Information Maximum Likelihood (FIML). This method does not require the imputation of any missing values. Whereas MI consists of imputation, analyses and pooling steps, FIML analyses the data in a single step. When the assumptions are met the two approaches should produce equivalent results. [REF] As FIML requires specialised software, not all analyses can be performed with standard software. [REF]

- Add table describing 2l- imputation models. like this https://bookdown.org/mwheymans/bookmi/multiple-imputation-models-for-multilevel-data.html

### Imputation process

#### Clustering specification
There are different strategies that can be adopted in the imputation process that account for clustering: inclusion of cluster indicator variable, performing a separate imputation process for each cluster, or performing a simultaneous imputation process by using an imputation method that accounts for clustering.(Stata: <https://www.stata.com/support/faqs/statistics/clustering-and-mi-impute/>) TODO: replace ref.

#### Selection of variables in the imputation model
- Inclusion of outcome variable [Moons] When the outcome is time-to-event, the Nelson-Aalen estimate of the time to event should be included as a covariate in the imputation model [REF]
- Imputation for interaction term, JAV, Machine lerarning methods (Random forest) and imputation separated by treatment group
- Prediction matrix specfication (table)

#### Convergency problems during imputation
- Check log 
- Problems of convergence on some 2l methods can be solved by use of overflux or quickpred. 
- Consider the use of scaled variables Vanburen. 
- Consider to swap imputation mehtod in problematic variables: The use of non parametrical imputation methods such as pmm. or downgrade to 1l methods.
- Also by the specification of post-processing in some variables can alivianate this. 
- Separated imputation process for each endpoint analysis

### Convergency 
Plots and also the prediction plots.

### Sensibility analysis
About the possibility of assumptions about Missing mechanism.

>>>>

### Focus on the Main Analysis
When imputing clustered data, we initially focus on the research question and the intended analysis, assuming the presence of non-incomplete values. This assessment sheds light on research hypotheses and the contemplated main(s) statistical model(s) and it provides insights into the data's structure (refer to the level table), variable types (e.g., confounders, auxiliary variables), and other considerations about variables relationships such as interactions or polynomial terms.

### Exploration of Available Data
Next, it is necessary to explore the information available in the data. Exploring individual variables using tools such as histograms or QQ plots can help to delineate variable distributions and plausible ranges of values and also identify input errors or outliers. 
Evaluating interactions between predictors using scatter plots or conditional plots, assessing collinearity using VIF or correlations, can help to glimpse nonlinear relationships that may affect the original formulation of the main model. Further testing of assumptions related to the response variable, including variance (homogeneity using conditional box plots), independence (e.g., ACF, variograms) and response vs. predictor relationships (conditional plots - x vs. y), may serve to choose an imputation model more suitable to the data.  In addition, the intraclass correlation coefficient (ICC) can be examined to assess cluster differences, aiding in the choice between the 2l and 1l methods for imputation.

One can also explore missingness, specifically missing patterns at the cluster level to identify types of patterns (e.g., non-monotonic, systematic) to guide the selection of an appropriate imputation approach in terms of computational efficiency (e.g., simpler regression imputation versus FCS in univariate patterns). Similarly, these missing patterns can also be used to identify potential predictors in individual imputation models by means of inflow criteria (quantifying the connections between missing data in one variable and other observed variables) and outflow criteria (determining the connections between observed values in one variable and missing data in other variables). 
Although several packages exist for visualizing missing data, this tutorial focuses on the recently implemented ggmice package.

### Assess Estimation Procedure Robustness to Missing Data
Prior to implementing multiple imputation, it is critical to assess the robustness of the estimation model when confronted with missing data. \cite{Alisson_2002} highlighted scenarios in which specific Maximum Likelihood (ML) estimation methods outperform Multiple Imputation (MI) methods.  Notably, when the response variable is the sole incomplete variable, mixed models demonstrate robustness to missing data under the Missing at Random (MAR) assumption and with a correct variance-covariance specification.\cite{Molenberghs_2007}. Furthermore, depending on the proportion of missingness (usually is <5%), opting for a simpler complete case analysis might be sufficient.

### Pre-imputation
Before proceeding with the imputation model, it is necessary to filter the dataset to include only the minimal variables required for the main model. If additional procedures are conducted during the analysis, it is essential to include variables associated with these procedures, such as confounders in the case of balancing techniques. Other relevant variables, like instrumental variables or auxiliary variables, may enhance parameter estimates even if not included in the main model, as they could be linked to the probability of missingness for some incomplete variables.

Furthermore, at the variable level, it is crucial to assess whether proxy variables should be considered or if deterministic imputation alone is adequate for imputing missing values. For instance, certain incomplete variables may not necessitate stochastic imputation methods like MICE. Instead, they can be effectively addressed through deductive imputation, where incomplete values are inferred from logical and deterministic relationships between variables. This approach is especially beneficial for variables that are functions of others, such as deriving a person's BMI from their weight and height. It also proves valuable for determining values for one-level variables from two-level ones, as seen in the context of Individual Participant Data (IPD). In such cases, incomplete information can be deduced from metadata, like inferring incomplete data about abortion in a country where abortion is illegal, or through cross-temporal or protocol-based deduction, such as imputing missing test values for deceased patients.

### Setting Imputation Model

##### Clustering Inclusion
Various strategies can be employed in the imputation process to account for clustering, including the inclusion of a cluster indicator variable, conducting a separate imputation process for each cluster, or employing a simultaneous imputation method that considers clustering (Stata: Clustering and MI Impute) TODO: Replace ref.

The choice of strategy depends primarily on the assumptions made in the main analysis and the constraints imposed by the analyzed data.Concerning data restrictions, if there are few clusters with many observations per cluster, the cluster indicator or imputation separated by groups may be appropriate \cite{Graham_2009}. Conversely, an imputation model that simultaneously imputes all clusters using a hierarchical model can be employed when there are more clusters or fewer observations per cluster \cite{Allison_2002}.

In a hierarchical imputation model, correlations between observations within clusters are modeled by a random effect, allowing estimation even with a limited number of observations per cluster. Multiple imputation models based on hierarchical models have been proposed, each with different assumptions.

<!--TODO: Add table 2l.imputation models -->

Before choosing a particular strategy, it is crucial to verify if the assumptions of the imputation model align with those of the main analysis. For example, using the cluster indicator strategy may introduce bias estimates when the model is based on a hierarchical structure \cite{Taaljard_2008}. Even if an imputation strategy congruent with the main model is preferred, it is essential to assess its appropriateness for the data, considering that less complex imputation strategies may lead to unbiased estimates in certain scenarios \cite(Bailey_2020).

#### Choice of Individual Imputation Methods
The initial step involves selecting the individual imputation model for each incomplete variable in the dataset. The mice package automatically proposes imputation methods based on the variable type for 1l variables. For 2l level variables, the micemd package's () function can be utilized. This function selects among the 2l imputation methods based on the size of the clusters and the proportion of missingness in each cluster.

In addition to the package-defined imputation methods, users can specify custom methods using the "I formula". This flexibility allows for the calculation of derived variables internally or adjustments to imputation methods based on specific conditions, such as conditioning the imputation model on the level of an incomplete covariate (e.g., pregnancy test on females).

#### Model Specification
The imputation model must be congenial with the main model \cite{Meng_1994}. Congeniality issues arise when the imputation model and the main model make different assumptions, often due to the omission of a polynomial or interaction term or the use of transformed variables.

The imputation model can include additional terms than the main model terms that do not lead to uncongeniality. For instance, it is advisable to include the outcome variable in the imputation model for prediction variables \cite{Moons}. In cases where the outcome is time-to-event, the Nelson-Aalen estimate of the time to the event should be included as a covariate in the imputation model [REF]. Furthermore, the inclusion of auxiliary variables, even if not part of the main model, can be associated with the probability of missingness, enhancing the likelihood of satisfying the Missing at Random (MAR) assumption and improving estimation efficiency \cite{Hardt_2012}.

The specification of imputation models is done on a variable basis, either using a prediction matrix where the type of prediction variable should be specified accordingly (see table) or through a list of formulas in the formula parameter. The latter option proves useful in formulating imputation models with polynomial terms or interactions compared to the prediction matrix specification, which requires the inclusion of additional terms as Just Another Variable (JAV).

<!--TODO: Add a table describing 2l-imputation models codes -->

However, even when considering an interaction term, depending on the adapted individual imputation model, it has been suggested, for instance, for treatment interaction effects, to conduct separated imputation by treatment group \cite{Zhang_2023}. Additionally, there are imputation models based on random forest or deep learning that can handle interaction and non-linear terms that do not require the explicit inclusion of the non-linear terms.

### Post-Imputation

During the imputation process, certain issues may arise that halt the process. In hierarchical model imputations, many issues are linked to methods based on parametric hierarchical models, which may struggle to estimate a model midway through the process. In such cases, it is advisable to examine the imputation log file to identify variables with imputation issues. Subsequently, reducing the number of predictors in the imputation model could be explored, either by using functions like quickpred or by considering variable transformations, such as scaling. Adjusting the level of the hierarchical model (e.g., using a homogenous variance model or a 1l model) can also be beneficial. Additionally, checking the range of imputed variables is crucial, as excessively large imputed values for one predictor could lead to convergence issues in other variables. This can be mitigated by including post-processing specifications on problematic variables or by using imputation models like Predicted Mean Matching (PMM), ensuring imputed values align with the observable values. In certain cases, a separate imputation strategy may be considered. For instance, in analyses involving multiple endpoints, conducting separate imputation processes for each endpoint might be preferable to a unified imputation process.

### Convergence and Sensitivity Analysis

Before commencing the analysis, it is imperative to verify the convergence of the imputation process. This is typically achieved through trace plots or by evaluating prediction plots.

While the majority of Multiple Imputation by Chained Equations (MICE) methods are based on Missing at Random (MAR) assumptions, field expert input may suggest that the Missing Not at Random (MNAR) mechanism could be plausible for certain variables. An MNAR variable is one in which the probability of missingness depends on an unobservable variable. This can occur when missingness is associated with the incomplete value itself (self-marking) or when there is an unobserved variable linked to both the value and the probability of missingness of the incomplete value (indirectly non-informative). Specifically, for the indirectly non-informative case in hierarchical datasets, imputation methods based on the Heckman method, such as those proposed by Hammon (2021) and Munoz (2023), can be considered.

### 2023-11-24 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
Update imputation workflow

\hypertarget{main-analysis}{%
\subsubsection{Main analysis}\label{main-analysis}}

When imputing clustered data, our first consideration should be the
research question and the type of analysis that researchers intend to
perform, assuming that non-incomplete values were present. This initial
assessment provides valuable insights into the data's structure (see the
level table), relationships among variables, and the causes of
missingness.

When multiple imputation is used to deal with missing data, as the
imputation and analysis process is performed separately, it is necessary
that imputation model being congenial with the main analysis model
(Meng, 1994), e.g.~if the main model accounts for the hierarchical
structure also imputation model should do it (Audigier, 2021).

\hypertarget{exploration-of-available-data}{%
\subsubsection{Exploration of available
data}\label{exploration-of-available-data}}

Then it is required to explore available information across clusters.
This can involve evaluating the Intra-class Correlation Coefficient
(ICC) to determine if a one-level imputation method is sufficient,
particularly if there are no substantial differences between clusters.
Additionally, it can be done by visualizing the data using available
package as \textbf{ggmice}, for instance examining density plots can
help us understand the variable type (normal, Poisson, categorical) and
its range of plausible values. Exploring relationships between
covariates through bivariate plots can be valuable in identifying
potential interactions or non-linearities.

In the context of missingness, we should assess patterns of systematic
and sporadic missing values and explore the relationships between the
missingness of an incomplete variable and observable data. This
evaluation allows us to consider the plausibility of Missing at Random
(MAR) mechanisms over Missing Completely at Random (MCAR) methods and
helps in selecting prediction variables to be included in the imputation
model (considering overflux).

\hypertarget{selecting-plausible-imputation-methods}{%
\subsubsection{Selecting plausible imputation
methods}\label{selecting-plausible-imputation-methods}}

Before to implement multiple imputation it is requiered to evaluate if
simpler imputation methods are suffice: - For example, certain
incomplete variables may not necessitate stochastic imputation methods
like MICE. Instead, they can be effectively addressed through deductive
imputation, where incomplete values can be inferred from logic and
deterministic relationships between variables. This approach is
particularly useful for variables that are functions of others, e.g.~a
person's BMI from their weight and height, determining values for
one-level variables from two-level ones, e.g.~in the context of
Individual Participant Data (IPD), incomplete information can be
extracted from metadata, such as deducing incomplete data about abortion
in a country where abortion is illegal, or engaging in cross-temporal or
protocol's deduction, like imputing missing test values for deceased
patients.

\begin{itemize}
\item
  Under monotonic pattern, the use of regression models
\item
  CC when the proportion of missing is minimal \textless5\%.
\item
  Analysis that are robust to missing data, for instance FIML models
  such as mixed model do not require imputation models when only the
  outcome variable is missing.An alternative approach to missing data is
  to use Full Information Maximum Likelihood (FIML). This method does
  not require the imputation of any missing values. Whereas MI consists
  of imputation, analyses and pooling steps, FIML analyses the data in a
  single step. When the assumptions are met the two approaches should
  produce equivalent results. {[}REF{]} As FIML requires specialised
  software, not all analyses can be performed with standard software.
  {[}REF{]}
\item
  Add table describing 2l- imputation models. like this
  \url{https://bookdown.org/mwheymans/bookmi/multiple-imputation-models-for-multilevel-data.html}
\end{itemize}

\hypertarget{imputation-process}{%
\subsubsection{Imputation process}\label{imputation-process}}

\hypertarget{clustering-specification}{%
\paragraph{Clustering specification}\label{clustering-specification}}

\hypertarget{selection-of-variables-in-the-imputation-model}{%
\paragraph{Selection of variables in the imputation
model}\label{selection-of-variables-in-the-imputation-model}}

\begin{itemize}
\tightlist
\item
  Inclusion of outcome variable {[}Moons{]} When the outcome is
  time-to-event, the Nelson-Aalen estimate of the time to event should
  be included as a covariate in the imputation model {[}REF{]}
\item
  Imputation for interaction term, JAV, Machine lerarning methods
  (Random forest) and imputation separated by treatment group
\item
  Prediction matrix specfication (table)
\end{itemize}

\hypertarget{convergency-problems-during-imputation}{%
\paragraph{Convergency problems during
imputation}\label{convergency-problems-during-imputation}}

\begin{itemize}
\tightlist
\item
  Check log
\item
  Problems of convergence on some 2l methods can be solved by use of
  overflux or quickpred.
\item
  Consider the use of scaled variables Vanburen.
\item
  Consider to swap imputation mehtod in problematic variables: The use
  of non parametrical imputation methods such as pmm. or downgrade to 1l
  methods.
\item
  Also by the specification of post-processing in some variables can
  alivianate this.
\item
  Separated imputation process for each endpoint analysis
\end{itemize}

\hypertarget{convergency}{%
\subsubsection{Convergency}\label{convergency}}

Plots and also the prediction plots.

\hypertarget{sensibility-analysis}{%
\subsubsection{Sensibility analysis}\label{sensibility-analysis}}

About the possibility of assumptions about Missing mechanism.

>>>>

\hypertarget{focus-on-the-main-analysis}{%
\subsubsection{Focus on the Main
Analysis}\label{focus-on-the-main-analysis}}

When imputing clustered data, we initially focus on the research
question and the intended analysis, assuming the presence of
non-incomplete values. This assessment sheds light on research
hypotheses and the contemplated main(s) statistical model(s) and it
provides insights into the data's structure (refer to the level table),
variable types (e.g., confounders, auxiliary variables), and other
considerations about variables relationships such as interactions or
polynomial terms.

\hypertarget{exploration-of-available-data}{%
\subsubsection{Exploration of Available
Data}\label{exploration-of-available-data}}

Next, it is necessary to explore the information available in the data.
Exploring individual variables using tools such as histograms or QQ
plots can help to delineate variable distributions and plausible ranges
of values and also identify input errors or outliers. Evaluating
interactions between predictors using scatter plots or conditional
plots, assessing collinearity using VIF or correlations, can help to
glimpse nonlinear relationships that may affect the original formulation
of the main model. Further testing of assumptions related to the
response variable, including variance (homogeneity using conditional box
plots), independence (e.g., ACF, variograms) and response vs.~predictor
relationships (conditional plots - x vs.~y), may serve to choose an
imputation model more suitable to the data. In addition, the intraclass
correlation coefficient (ICC) can be examined to assess cluster
differences, aiding in the choice between the 2l and 1l methods for
imputation.

One can also explore missingness, specifically missing patterns at the
cluster level to identify types of patterns (e.g., non-monotonic,
systematic) to guide the selection of an appropriate imputation approach
in terms of computational efficiency (e.g., simpler regression
imputation versus FCS in univariate patterns). Similarly, these missing
patterns can also be used to identify potential predictors in individual
imputation models by means of inflow criteria (quantifying the
connections between missing data in one variable and other observed
variables) and outflow criteria (determining the connections between
observed values in one variable and missing data in other variables).
Although several packages exist for visualizing missing data, this
tutorial focuses on the recently implemented ggmice package.

\hypertarget{assess-estimation-procedure-robustness-to-missing-data}{%
\subsubsection{Assess Estimation Procedure Robustness to Missing
Data}\label{assess-estimation-procedure-robustness-to-missing-data}}

Prior to implementing multiple imputation, it is critical to assess the
robustness of the estimation model when confronted with missing data.
\cite{Alisson_2002} highlighted scenarios in which specific Maximum
Likelihood (ML) estimation methods outperform Multiple Imputation (MI)
methods. Notably, when the response variable is the sole incomplete
variable, mixed models demonstrate robustness to missing data under the
Missing at Random (MAR) assumption and with a correct
variance-covariance specification.\cite{Molenberghs_2007}. Furthermore,
depending on the proportion of missingness (usually is \textless5\%),
opting for a simpler complete case analysis might be sufficient.

\hypertarget{pre-imputation}{%
\subsubsection{Pre-imputation}\label{pre-imputation}}

Before proceeding with the imputation model, it is necessary to filter
the dataset to include only the minimal variables required for the main
model. If additional procedures are conducted during the analysis, it is
essential to include variables associated with these procedures, such as
confounders in the case of balancing techniques. Other relevant
variables, like instrumental variables or auxiliary variables, may
enhance parameter estimates even if not included in the main model, as
they could be linked to the probability of missingness for some
incomplete variables.

Furthermore, at the variable level, it is crucial to assess whether
proxy variables should be considered or if deterministic imputation
alone is adequate for imputing missing values. For instance, certain
incomplete variables may not necessitate stochastic imputation methods
like MICE. Instead, they can be effectively addressed through deductive
imputation, where incomplete values are inferred from logical and
deterministic relationships between variables. This approach is
especially beneficial for variables that are functions of others, such
as deriving a person's BMI from their weight and height. It also proves
valuable for determining values for one-level variables from two-level
ones, as seen in the context of Individual Participant Data (IPD). In
such cases, incomplete information can be deduced from metadata, like
inferring incomplete data about abortion in a country where abortion is
illegal, or through cross-temporal or protocol-based deduction, such as
imputing missing test values for deceased patients.

\hypertarget{setting-imputation-model}{%
\subsubsection{Setting Imputation
Model}\label{setting-imputation-model}}

\hypertarget{clustering-inclusion}{%
\subparagraph{Clustering Inclusion}\label{clustering-inclusion}}

Various strategies can be employed in the imputation process to account
for clustering, including the inclusion of a cluster indicator variable,
conducting a separate imputation process for each cluster, or employing
a simultaneous imputation method that considers clustering (Stata:
Clustering and MI Impute) TODO: Replace ref.

The choice of strategy depends primarily on the assumptions made in the
main analysis and the constraints imposed by the analyzed
data.Concerning data restrictions, if there are few clusters with many
observations per cluster, the cluster indicator or imputation separated
by groups may be appropriate \cite{Graham_2009}. Conversely, an
imputation model that simultaneously imputes all clusters using a
hierarchical model can be employed when there are more clusters or fewer
observations per cluster \cite{Allison_2002}.

In a hierarchical imputation model, correlations between observations
within clusters are modeled by a random effect, allowing estimation even
with a limited number of observations per cluster. Multiple imputation
models based on hierarchical models have been proposed, each with
different assumptions.

Before choosing a particular strategy, it is crucial to verify if the
assumptions of the imputation model align with those of the main
analysis. For example, using the cluster indicator strategy may
introduce bias estimates when the model is based on a hierarchical
structure \cite{Taaljard_2008}. Even if an imputation strategy congruent
with the main model is preferred, it is essential to assess its
appropriateness for the data, considering that less complex imputation
strategies may lead to unbiased estimates in certain scenarios
\textbackslash cite(Bailey\_2020).

\hypertarget{choice-of-individual-imputation-methods}{%
\paragraph{Choice of Individual Imputation
Methods}\label{choice-of-individual-imputation-methods}}

The initial step involves selecting the individual imputation model for
each incomplete variable in the dataset. The mice package automatically
proposes imputation methods based on the variable type for 1l variables.
For 2l level variables, the micemd package's () function can be
utilized. This function selects among the 2l imputation methods based on
the size of the clusters and the proportion of missingness in each
cluster.

In addition to the package-defined imputation methods, users can specify
custom methods using the ``I formula''. This flexibility allows for the
calculation of derived variables internally or adjustments to imputation
methods based on specific conditions, such as conditioning the
imputation model on the level of an incomplete covariate (e.g.,
pregnancy test on females).

\hypertarget{model-specification}{%
\paragraph{Model Specification}\label{model-specification}}

The imputation model must be congenial with the main model
\cite{Meng_1994}. Congeniality issues arise when the imputation model
and the main model make different assumptions, often due to the omission
of a polynomial or interaction term or the use of transformed variables.

The imputation model can include additional terms than the main model
terms that do not lead to uncongeniality. For instance, it is advisable
to include the outcome variable in the imputation model for prediction
variables \cite{Moons}. In cases where the outcome is time-to-event, the
Nelson-Aalen estimate of the time to the event should be included as a
covariate in the imputation model {[}REF{]}. Furthermore, the inclusion
of auxiliary variables, even if not part of the main model, can be
associated with the probability of missingness, enhancing the likelihood
of satisfying the Missing at Random (MAR) assumption and improving
estimation efficiency \cite{Hardt_2012}.

The specification of imputation models is done on a variable basis,
either using a prediction matrix where the type of prediction variable
should be specified accordingly (see table) or through a list of
formulas in the formula parameter. The latter option proves useful in
formulating imputation models with polynomial terms or interactions
compared to the prediction matrix specification, which requires the
inclusion of additional terms as Just Another Variable (JAV).

However, even when considering an interaction term, depending on the
adapted individual imputation model, it has been suggested, for
instance, for treatment interaction effects, to conduct separated
imputation by treatment group \cite{Zhang_2023}. Additionally, there are
imputation models based on random forest or deep learning that can
handle interaction and non-linear terms that do not require the explicit
inclusion of the non-linear terms.

\hypertarget{post-imputation}{%
\subsubsection{Post-Imputation}\label{post-imputation}}

During the imputation process, certain issues may arise that halt the
process. In hierarchical model imputations, many issues are linked to
methods based on parametric hierarchical models, which may struggle to
estimate a model midway through the process. In such cases, it is
advisable to examine the imputation log file to identify variables with
imputation issues. Subsequently, reducing the number of predictors in
the imputation model could be explored, either by using functions like
quickpred or by considering variable transformations, such as scaling.
Adjusting the level of the hierarchical model (e.g., using a homogenous
variance model or a 1l model) can also be beneficial. Additionally,
checking the range of imputed variables is crucial, as excessively large
imputed values for one predictor could lead to convergence issues in
other variables. This can be mitigated by including post-processing
specifications on problematic variables or by using imputation models
like Predicted Mean Matching (PMM), ensuring imputed values align with
the observable values. In certain cases, a separate imputation strategy
may be considered. For instance, in analyses involving multiple
endpoints, conducting separate imputation processes for each endpoint
might be preferable to a unified imputation process.

\hypertarget{convergence-and-sensitivity-analysis}{%
\subsubsection{Convergence and Sensitivity
Analysis}\label{convergence-and-sensitivity-analysis}}

Before commencing the analysis, it is imperative to verify the
convergence of the imputation process. This is typically achieved
through trace plots or by evaluating prediction plots.

While the majority of Multiple Imputation by Chained Equations (MICE)
methods are based on Missing at Random (MAR) assumptions, field expert
input may suggest that the Missing Not at Random (MNAR) mechanism could
be plausible for certain variables. An MNAR variable is one in which the
probability of missingness depends on an unobservable variable. This can
occur when missingness is associated with the incomplete value itself
(self-marking) or when there is an unobserved variable linked to both
the value and the probability of missingness of the incomplete value
(indirectly non-informative). Specifically, for the indirectly
non-informative case in hierarchical datasets, imputation methods based
on the Heckman method, such as those proposed by Hammon (2021) and Munoz
(2023), can be considered.
